In-silico molecular enrichment and clearance of the human intracranial space

### 2024-12-20 paper/supplementary_main.tex
Tidy up supplementary PVS peristalsis.

    A & Baseline with sheaths & $R_2 = 2 R_1$ & $\xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\ 

@@

  \caption{Overview of models (-- denotes as immediately above). Each

@@

\mer{MER: This needs a pass to update to new notation and new style.}

We use the theoretical framework introduced by Gjerde et al.~\cite{gjerde2023directional} to compute an analytic estimate of the time-averaged (or \emph{net}) flow rates $\langle Q'_i \rangle$ (mm$^3$/s) induced by peristaltic pumping in a perivascular network$\Lambda = \cup_i \Lambda_i$. The motion of the (inner) vascular wall is described by a periodic traveling (peristaltic) wave of relative amplitude $\varepsilon$, wave length $\lambda$ (mm) and frequency $f$, acting normal to the wall. By definition, the wave number is $k = 2 \pi/\lambda$ and the angular frequency is $\omega = 2 \pi f$. Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L$, baseline inner radius $r_{o_i}'$, fixed outer radius $r_{e_i}'$, and outer-to-inner ratio $\beta_i = r_{e_i}'/r_{o_i}'$ (see also \Cref{tab:pvs:parameters}). These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances $\mathcal{R}_{o, i}$ and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics. Since the analytical estimate is derived under the assumption that $k L \approx \mathcal{O} 1)$~\cite{gjerde2023directional} and has been verified against numerical simulations for $k L$ of the order $10^{-1}-10^2$~\cite[Table I]{gjerde2023directional}, we consider it applicable for the wave lengths and vascular network data considered here in which $k L$ ranges from \mer{$X$ to $Y$.}

\begin{table}
  \small
  \begin{tabular}{llrc}
    \toprule
    Parameter & Description & Value(s)  & Ref.\\ 
    \midrule
    Relative PVS size $\beta$ & $\beta = \beta_i = r_{e_i} / r_{o_i}$ & 2 & $\approx$\cite{eide2024functional} \\
    Wave frequency $f$ & Traveling wave frequency of vascular motion & $0.1-1.0$ Hz & \discuss{\cite{gjerde2023directional}} \\
    Wave length $\lambda$ & Traveling wave length of vascular motion & $20-2000$ mm & \discuss{\cite{broggini2024long, gjerde2023directional}} \\
    Wave amplitude & Relative amplitude of inner wall motion & $1-10\%$ & \discuss{\cite{gjerde2023directional}} \\
    Vascular radii $\{r_{o_i}\}$ & Vascular network radii data (avg $\pm$ std, $\min$, $\max$) & $X$, $X$, $X$  mm & \cite{hodneland2019new} \\
    Vascular branch lengths $\{L_{i}\}$ & Path length between vasc. junctions (avg $\pm$ std, $\min$, $\max$) & $X \pm Y$, $X$, $X$  mm & \discuss{--} \\
    \bottomrule
  \end{tabular}
  \caption{Perivascular flow induced by vascular wall motion: parameters and key data characteristics.}
  \label{tab:pvs:parameters}
\end{table}
% Vascular radii r_o (min, max, avg, std) (mm):  ....
% Vascular branch lengths L (avg, std, min, max) (mm): ...

\input{tikz/fig_pvs_flow}
This theoretical formalism~\cite{gjerde2023directional} is defined
relative to a network in the form of a directed, bifurcating tree with
a single supply node/root $i_0$. To extend to a network of cerebral
arteries with multiple supply nodes (such as the basilar and two
internal carotid arteries in the current data
set~\cite{hodneland2019new}), we separate the network $\Lambda$ into
edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
each of the supply nodes (\Cref{fig:app:peristalsis}). Each node is assigned to the subnetwork associated with the nearest supply node, and edges between nodes are preserved. Next, we compute a minimal, bifurcating and directed tree representation of each subnetwork: $\mathcal{T}^1, \mathcal{T}^2, \mathcal{T}^3$. Each tree $\mathcal{T}^j = \cup E_n^j$ consists of the subset of the nodes from $\Lambda^j$ that have degree 1 (are leaf or root nodes) or degree $3$ (are true bifurcation points), and each path between nodes with degree $\not = 2$ in $\Lambda^j$ is represented in $\mathcal{T}^j$ by an edge $E^j$ with edge length $L$ corresponding to the total length of the original path and edge radius $r_o$ as the average of the path radii. For the few (2) nodes in $\Lambda$ with degree $>3$; i.e.~junctions where a branch splits into more than two downstream branches, only the two longest downstream paths are included in $\mathcal{T}^j$.

For each subtree $\mathcal{T}^j = \cup_n E_n^j$, we compute the
time-averaged downstream flow rate $\langle Q'_n \rangle$ induced by
the vascular wall motion for each edge $n$ via~\cite[eq.~(5),
  (34)]{gjerde2023directional}. We next uniquely assign the flow rate
$\langle Q'_n \rangle$ to each of the branches $\Lambda_i$ that form
the path $E_n^j$, thus yielding $\langle Q'_i \rangle$ for each
perivascular segment $\Lambda_i$ (except those ignored above, for
which we set a flow rate of zero). Finally, we define the mean
longitudinal perivascular velocity $\langle u^{\rm PVS}_i \rangle$ by
dividing the flow rate $\langle Q'_i \rangle$ by the cross-section
area $A_i = \pi (r_{e_i}^2 - r_{o_i}^2) = \pi (\beta_i^2 - 1) r_{o_i}^2$

>>>>

    A & Baseline & $R_2 = 2 R_1$ & $\xi$ & $10^{-4}$ mm$^2$/s\cite{hornkjol2022csf} & $\phi R D^{\rm gad}_{\rm eff}$\cite{sykova2008diffusion, valnes2020apparent} & $\hat{u}_{\rm prod} + \hat{u}_{\rm cardiac}$ & $\mathbf{u}_{\rm prod}$ & 0 \\ 

@@

  \caption{Overview of models. (-- denotes as immediately above). Each

@@

We use the theoretical framework previously introduced by Gjerde et al.~\cite{gjerde2023directional} to compute an analytic estimate of the time-averaged (or \emph{net}) flow rates $\langle Q'_i \rangle$ (mm$^3$/s) induced by peristaltic pumping in a perivascular network $\Lambda = \cup_i \Lambda_i$. The motion of the (inner) vascular wall is described by a periodic traveling (peristaltic) wave of relative amplitude $\varepsilon$, wave length $\lambda$ (mm) and frequency $f$, acting normal to the wall. By definition, the wave number is $k = 2 \pi/\lambda$ and the angular frequency is $\omega = 2 \pi f$. Each PVS segment $\Lambda_i$ has length $L_i$ with wave-relative length $\ell_i = k L$, baseline inner radius $R_1^i$, fixed outer radius $R_2^i$, and outer-to-inner ratio $\beta_i = R_2^i/R_1^i$. These geometric parameters and the assumption of annular cylindrical PVS segments yield hydraulic resistances $\mathcal{R}_{o, i}$ and additional characteristic parameters, see \cite{gjerde2023directional} for the complete definitions and schematics. Since the analytical estimate is derived under the assumption that $k L \approx \mathcal{O} (1)$~\cite{gjerde2023directional} and has been verified against numerical simulations for $k L$ of the order $10^{-1}-10^2$~\cite[Table I]{gjerde2023directional}, we consider it applicable for the wave lengths and vascular network data considered here in which $k L$ range from 0.15 to 30 for the strong vasomotion and 0.0015 to 0.30 for the cardiac waves (\Cref{tab:pvs:parameters}).
\begin{table}
  \small
  \begin{tabular}{llrc}
    \toprule
    Parameter & Description & Value(s)  & Ref.\\ 
    \midrule
    Relative PVS size $\beta$ & $\beta = \beta_i = R_2^i / R_1^i$ & 2 & $\approx$\cite{eide2024functional} \\
    Wave frequency $f$ & Traveling wave frequency of vascular motion & $0.1-1.0$ Hz & \discuss{\cite{gjerde2023directional}} \\
    Wave length $\lambda$ & Traveling wave length of vascular motion & $0.02-2$ mm & \discuss{\cite{broggini2024long, gjerde2023directional}} \\
    Wave amplitude & Relative amplitude of inner wall motion & $1-10\%$ & \discuss{\cite{gjerde2023directional}} \\
    \midrule
    Cardiac frequency $f_c$ & Frequency of human cardiac pulse wave  & $1.0$ Hz & \cite{jung2021novel} \\
    Cardiac wave length $\lambda_c$ & Wave length of human cardiac pulse wave & $2.0$m & \\
    Cardiac amplitude $\varepsilon_c$ & Wall displacement amplitude of cardiac pulse wave & 1\%  & \\
    \midrule
    Vasomotion frequency $f_v$ & Frequency of slow vasomotion wave  & $0.1$ Hz & \cite{jung2021novel} \\
    Vasomotion wave length $\lambda_v$ & Wave length of vasomotion wave & $0.02$ m & \cite{broggini2024long, gjerde2023directional} \\
    Vasomotion amplitude $\varepsilon_v$ & Wall displacement amplitude of vasomotion wave & 10\%  & \cite{broggini2024long, gjerde2023directional}\\
    %Vascular radii $\{r_{o_i}\}$ & Vascular network radii data (avg $\pm$ std, $\min$, $\max$) & $X$, $X$, $X$  mm & \cite{hodneland2019new} \\
    %Vascular branch lengths $\{L_{i}\}$ & Path length between vasc. junctions (avg $\pm$ std, $\min$, $\max$) & $X \pm Y$, $X$, $X$  mm & \discuss{--} \\
    \bottomrule
  \end{tabular}
  \caption{Perivascular flow induced by vascular wall motion: overview of parameters.}
  \label{tab:pvs:parameters}
\end{table}
% Vascular radii r_o (min, max, avg, std) (mm):  ....
% Vascular branch lengths L (avg, std, min, max) (mm): ...

\input{tikz/fig_pvs_flow}

This theoretical formalism~\cite{gjerde2023directional} is defined
relative to a network in the form of a directed, bifurcating tree with
a single supply node/root $i_0$. To extend to a network of cerebral
arteries with multiple supply nodes (such as the basilar and two
internal carotid arteries in the current data
set~\cite{hodneland2019new}), we separate the network $\Lambda$ into
edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
each of the supply nodes (\Cref{fig:results1}A, C,
\Cref{fig:app:peristalsis}). Each node is assigned to the subnetwork
associated with the nearest supply node, and edges between nodes are
preserved. Next, we compute a minimal, bifurcating and directed tree
representation of each subnetwork: $\mathcal{T}^1, \mathcal{T}^2,
\mathcal{T}^3$. Each tree $\mathcal{T}^j = \cup E_n^j$ consists of the
subset of the nodes from $\Lambda^j$ that have degree 1 (are leaf or
root nodes) or degree $3$ (are true bifurcation points), and each path
between nodes with degree $\not = 2$ in $\Lambda^j$ is represented in
$\mathcal{T}^j$ by an edge $E^j$ with edge length $L$ corresponding to
the total length of the original path and edge radius $R^j$ as the
average of the path radii. \discuss{For the few (2) nodes in $\Lambda$ with
degree $>3$; i.e.~junctions where a branch splits into more than two
downstream branches, only the two longest downstream paths are
included in $\mathcal{T}^j$.}

For each subtree $\mathcal{T}^j = \cup_n E_n^j$, we compute the
time-averaged downstream flow rate $\langle Q'_n \rangle$ induced by
the vascular wall motion for each edge $n$ via~\cite[eq.~(5),
  (34)]{gjerde2023directional}. We next uniquely assign the flow rate
$\langle Q'_n \rangle$ to each of the branches $\Lambda_i$ that form
the path $E_n^j$, thus yielding $\langle Q'_i \rangle$ for each
perivascular segment $\Lambda_i$ (except those ignored above, for
which we set a flow rate of zero). Finally, we define the mean
longitudinal perivascular velocity induced by the peristaltic wave
$\langle \hat{u}^x_i \rangle$ by dividing the flow rate $\langle Q'_i
\rangle$ by the cross-section area $A_i = \pi ({R_2^i}^2 - {R_1^i}^2)
= \pi (\beta_i^2 - 1) (R_1^i)^2$.
